\textbf{Access Control for Content Distribution Network Assets}

### 2016-06-13 tex/paper.tex
Added half of the implementation fixed

\subsubsection{Problem description}
The current system does not offer proper security measurements, is lacking a lot of features that
are needed and does not scale very well, hence a new system are to be developed. This work is about
examining a way of implementing Copy-on-Write in a high-level system like this, which should solve
the scalability problem and make it possible to implement wanted features like snapshots, cloning
and concurrent modifications of content.

@@

To clarify what is meant by Perius being high-level software the stack in Figure~\ref{fig:stack} can
be examined. Theoretically the Copy-on-Write features could be implemented in any of the separate
parts of the stack to support the features that Perius has to achieve, in this work only the second
highest layer is considered, the Perius back-end.

@@

\item[Javascript]
is a programming language mainly focussed on running code in web browsers~\cite{JAVASCRIPT}. In this
implementation the ECMAScript 6 (ES6) language specification is used, which adds a lot of syntactic
sugar compared to what usually is called normal JavaScript. An example of such syntactic sugar is
lambda expressions. React and Reflux are both built with JavaScript in mind and is used on top of
JavaScript is this implementation, as can be seen in Figure~\ref{fig:tech_stack}.

\item[SuperAgent]
is a small HTTP request library which can be used from client-side Javascript or as a node.js
module~\cite{SUPERAGENT}. It is used to form the requests to the REST API from the front-end, which
can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).

@@

\item[Reflux] 
is an idea and a simple library of how to structure your application~\cite{REFLUX}. It features a
unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example
Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used together with
React for this implementation to define a structure of the flow of data and state changes.

@@

\item[Scala]
is a multi-paradigm programming language. It most commonly runs on the JVM and unlike Java it
supports most functional programming features at the same time as it supports object-oriented
programming~\cite{SCALA}. Scala is used to program the back-end and the Spray toolkit is used with
it to simplify the API.

@@

In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
communication, which can be seen in Figure~\ref{fig:tech_stack} (Arrow 1 and 2).

@@

As can be seen several expected endpoints are missing, this is intentional as the operations missing
can be performed in a more efficient way. Such endpoint is for example \textit{GET 
projects/\{id\}/containers} as all containers exist in \textit{GET projects/\{id\}} and the interface
should present a file structure where both content and containers are shown.

% \hfill \\ %\label{sec:spray}
\item[Spray]~\cite{CASBAH} % Spray in casbah book
is an open-source toolkit written in Scala and for Scala. It is used to make it easier to build
REST/HTTP APIs and those abilities is what it is used in the implementation.

\item[MongoDB]% \hfill \\ %\label{sec:spray}
is a document-oriented database, which means that it does not have the concept of rows as
normal relational databases has. Instead each entity in the database is stored as a document which
is not fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks the support for joins to
improve its possibility to scale, which can be a big downside to some applications containing the
need for such logic. MongoDB is used as a persistent storage for the back-end to store all the
projects and nodes in the tree structures and also the files uploaded to the system. The Scala logic
code uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB, which can be seen in
Figure~\ref{fig:tech_stack} (Arrow 4 and 5).

\par MongoDB was chosen as the persistent storage because of its quick lookups and because of its
internal storage format called BSON, which is very similar to JSON which the API is using. As the
formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
scale in the future it is very easy to distribute MongoDB and, if needed, the system can easily be
migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
and can therefore make the system scale even further~\cite{REACTIVEMONGO}.

\par All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files 
according to the size limit of MongoDB objects, which is currently 4MB. The advantage of this is 
that backups of the Perius system is easily done through a database backup, no separate files 
needs to be backed up. Another advantage that is given by this is that you can retrieve specific 
ranges of a file, although that advantage is not needed in the Perius implementation. For
scalability this could be used to retrieve different parts of a file from different servers, 
normal load balancing would probably work better in a system like Perius where no extremely large 
files are expected to be stored.

@@

\end{description}
\subsection{Methods for determining\\implementation details}
This chapter introduces the different methods used to determine how the new system should be
implemented, which persistent storage it should use and how the estimation of long-term scaling 
was done.
\\
\par As this work was not mainly about the system architecture, but rather about evaluating the
effects of using Copy-on-Write, the decisions taken for the different technologies used were simply
made from what the author was comfortable with and knew would have little impact on the evaluation. 
With that said there are no comparisons of for example different databases, protocols or API
solutions in this paper.

@@

\par A low timeout was set to make sure that the system could handle operations faster than they
queued up. This was to be certain that the system could sustain the number of active users during a
longer time frame without having a queue which grows beyond a manageable size.

@@

\subsubsection{Snapshot Functionality}
The Copy-on-Write concept is not only well-suited for consistency requirements, it is also works 
well for creating snapshots.

@@

\subsubsection{Examples of Copy-on-Write system implementations}
\subsubsubsection{BTRFS} \label{sec:btrfs}
Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
tree can refer to the same node each node keeps track of how many parents it has by a reference
counter so that the node can be deallocated once the node does not have any parents any more. The
reference counter is not stored in the nodes themselves but rather in a separate data structure so
that a node's counter can be modified without modifying the node itself and therefore eludes the
Copy-on-Write that would have to occur.

@@

\subsubsubsection{Mach kernel}
In the mid 80's when the development of the Mach kernel started, there was problems with that
physically copying memory was too slow. To minimise the copying of memory, Copy-on-Write was
implemented. It was implemented so that virtual copy operations could be done and so that tasks
could share read-write memory~\cite{MACH}.

\subsubsection{Perius}
As the persistent storage, used in this implementation (Section~\ref{sec:tech}), does
not implement transactions or locks a lot of different problems can occur when several clients are
working on the same data set at the same time. Such problems could be race conditions or determining
the happened-before relation~\cite{LAMPORT}. In this work the race condition problem is solved by 
implementing Copy-on-Write on data with high consistency requirements. The happened-before relation
is not fully solved, but suggestions of how to solve it is given in Section~\ref{sec:conflict_res}.

@@

\subsubsection{Perius}
Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
back-end written in Scala and a front-end written in Javascript (ES6), but they are both
interchangeable. The back-end has a REST API running, which is how the front-end communicates with
the back-end.

@@

The service features a virtual file structure over the assets and metadata that have been stored,
as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
whole containers as well as individual files, audit and auth logging, multiple separate project 
and a design which allows for an interchangeable persistent storage.

@@

\subsubsection{Entities of the Implementation}
This section explains the different types of entities, and their interaction, that are handled by
Perius. They are explained similarly to how they are explained for the model in
Section~\ref{sec:elements}, but with more implementation specific details.

@@

In the user interface the stored data is always represented and worked on as a tree, or a forest if
all projects are considered. For a very small project that tree could look like the example in
Figure~\ref{fig:project_tree}. This is however not what the actual storage of the data looks like.

@@

The actual storage of the data is represented by documents in a flat document storage, which is in
this case MongoDB. A very simplified version of how the data is stored can be seen in
Figure~\ref{fig:storage}, however in reality it is not quite so table-like. One important point that
differs in the example and in the actual storage is that GUIDs are used for identifiers and not
incremented numbers, as that would not work in distributed environments.

The data is not stored as a tree (see Example~\ref{fig:storage}), but rather as separate documents 
loosely referring to each other. How the different operations perform their actions and what 
implications these actions have on the operations run-time complexities are not obvious and that 
is what is discussed in this section.\\

\par \textbf{GET single entity} \\
When the back-end is handed a GET request it looks at what type of entity that is requested and
returns the document with the correct ID from the corresponding collection. As the collections are
hash indexed on ID this operation can be made in constant time, $O(1)$ and thus $\theta(1)$. \\

\par \textbf{GET project tree} \\
When requesting a project the full project tree with all entities except files has to be built as
the entities not are stored as a tree in MongoDB. The tree is built layer by layer by fetching
content with the same parent ID, starting from the root of the project. Parent ID's are also indexed
by a hash index which makes it possible to get all nodes with the same parent ID in constant time.
This has to be done for every level with a different parent. The worst case for building the tree is
thus $\theta(h)$ where h is the height of the tree and all of the subtrees only have one node each.
The average case can not be calculated as the tree's balance is not determined by an algorithm but
rather by a user, but from observation one could argue that the trees usually have several nodes for
each parent node in the tree, which then would give a logarithmic average case, $\theta(log(n))$.\\

\par \textbf{POST/Create entity} \\
Creating a new entity in the system is a constant time operation for containers and content as they
only have to be inserted as new documents in MongoDB, which is a $O(1)$
operation~\cite{MONGOPERFORMANCE}. Even though the insertion itself has constant run-time the index
needs to be updated or sometimes even rebuilt, which can slow down the time before the document is
accessible in constant time. When inserting files they need to be chunked into pieces and inserted
into GridFS, this results in $c/s$ insertions where $c$ is the size of the file and s is the chunk size.
The insertion of files can then be regarded as $\theta(c)$ if the size of the file is regarded, but
if $c$ is regarded as the number of files in the system the upper bound is constant as the insertion
for files is made in the same way as for other documents.\\

\par \textbf{Snapshot creation} \\
In the creation of a snapshot all metadata in the tree that the snapshot is taken of is duplicated,
meanwhile the files remain the same, this results in $n$ insertions where $n$ is the number of
containers and content entities in the subtree. A snapshots insertion runtime is therefore
always $\theta(n)$. \\

\par \textbf{Delete entity} \\ 
Deleting an entity is a constant time operation when that entity does not have any children. When
the entity has children all those children has to be deleted too which will be done in the same
manner as the project tree is built and therefore results in the same runtime, $\theta(n)$ as worst
case and $\theta(log(n))$ as average (with the same argumentation as in \textit{GET project tree}),
where $n$ is the number of entities in the subtree.\\

@@

\subsubsection{GET requests} \label{sec:GET_REQUESTS}
To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
requests} are performed. Most responses are JSON structures, except for the ones in
Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}, which can be seen in the Response data part
of the listings.\\

@@

\par
The first request (Listing~\ref{OKREQUEST}) tries to maximise the number of requests that spray-can
(HTTP server) can handle with the given specifications and therefore the server just returns 200
(OK).

@@

\par
The second request (Listing~\ref{DBREQUEST}) performs the simplest API call which involves getting
all currently stored projects from the MongoDB instance. This request does not yield any extra
computation in the back-end, apart from converting the MongoDB format to the API's JSON result
format.

@@

\par
As the second request (Listing~\ref{DBREQUEST}) involves the server answering with some payload that
is generated from the persistent storage, the third request (Listing~\ref{STATICREQUEST}) is simply
the same amount of static payload as the second request but without involving any connections to
MongoDB. This was done to be able to estimate what impact the round trip time to MongoDB and the
conversions to the response format had on the throughput.

@@

The result in Listing~\ref{TREEREQUEST} shows an average of 40 requests/s, which means this is the
real bottleneck in the application. There are two solutions to fix this bottleneck, as the project
tree will not be modified nearly as often as it is requested, the first solution would be to cache
the results of the tree requests and invalidate those caches when the tree is modified. The second
solution would be to make a more computationally feasible way of building and fetching the whole
tree from the database. This is left to implement if the need comes for it in the future. Even
though this massive bottleneck is present, it will not be a problem in a small scale production
environment as the tree is fetched approximately once every 30 seconds by active users, which means
that the application still could support at least 1200 very active users.

@@

\par
With caching turned on, on a single back-end, the application can theoretically support several 
million active users which are not posting any content, this is simulated in 
Listing~\ref{CACHEREQUEST}.

@@

\par As a project tree is built by querying documents parent ID an optimisation that could be done was
to create an index of the documents parent IDs. This drastically improved the size of the tree that
the server was able to handle within the time-out, going from being able to handle 4000 documents in 
one tree to at least 50000 documents, which the browser instead will have problems displaying, 
represented in the UI, without lag.

\par After adding indices, the building of the project tree was able to finish 50 times faster and
the server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
maximum number of active clients that one node can handle will be >60000, without taking the web
servers limits or content posting into consideration.

>>>>

\subsubsection{Problem description}
The current system does not offer proper security measurements as its mechanism for keeping private
content out of the hands of the public is based on security by obscurity, which is further explained
in Section~\ref{sec:current_system}.

It is also lacking several features that are needed and does not scale very well, hence a new system
is to be developed. Features that needs to be implemented are for example snapshots, cloning and
proper support of concurrent modifications of content.

This work is about examining a way of implementing Copy-on-Write in a high-level system, which
should solve the scalability problem and make it possible to implement wanted features.

@@

To clarify what is meant by Perius being high-level software the stack in Figure~\ref{fig:stack} can
be examined. Theoretically the Copy-on-Write features could be implemented in any part of the stack,
in this work only the second highest layer is considered, the Perius back-end.

@@

\item[Javascript]
is a programming language mainly focussed on running code in web browsers~\cite{JAVASCRIPT}. In this
implementation the ECMAScript 6 (ES6) language specification is used, which adds a lot of syntactic
sugar compared to what usually is called normal JavaScript. An example of such syntactic sugar is
lambda expressions. React and Reflux are both built with JavaScript in mind and is used on top of
JavaScript in this implementation.

\item[SuperAgent]
is a small HTTP request library which can be used from client-side Javascript or as a node.js
module~\cite{SUPERAGENT}. It is used to form the requests to the REST API from the front-end, arrow
1 and 2 in Figure~\ref{fig:tech_stack}.

@@

\item[Reflux] 
is an idea and a simple library of how to structure your application~\cite{REFLUX}. It features a
unidirectional dataflow (see Figure~\ref{fig:reflux}) which makes it more suitable, than for example
Flux~\cite{FLUX}, when using a functional reactive programming style. Reflux is used together with
React to define a structure of the flow of data and state changes.

@@

\item[Scala]
is a multi-paradigm programming language. It most commonly runs on the JVM and unlike Java it
supports most functional programming features at the same time as it supports object-oriented
programming~\cite{SCALA}. Scala is used to program the back-end and the Spray toolkit is used to
simplify the API.

@@

In this work a RESTful API was implemented and used for back-end $\Iff$ front-end
communication, arrow 1 and 2 in Figure~\ref{fig:tech_stack}.

@@

As can be seen several expected endpoints are missing, this is intentional as the operations missing
can be performed in a more efficient way. Such endpoint is for example GET 
projects/\{id\}/containers as all containers exist in GET projects/\{id\} and the interface
should present a file structure where both content and containers are shown.

% \hfill \\ %\label{sec:spray}
\item[Spray]~\cite{CASBAH} % Spray in casbah book
is an open-source toolkit written in Scala and for Scala. Spray is mainly used to make it easier to
build REST/HTTP APIs.

\item[Casbah]~\cite{CASBAH} % Spray in casbah book
is a Scala toolkit and a java driver wrapper used to interface with MongoDB instances in a more
Scala oriented way.

\item[MongoDB]% \hfill \\ %\label{sec:spray}
is a document-oriented database, which means that it does not have the concept of rows as normal
relational databases has. Instead each entity in the database is stored as a document that is not
fixed to a predefined table structure~\cite{MONGODB}. MongoDB lacks support for join operations to
improve its possibility to scale, which can be a big disadvantage for applications having the need
for such logic. MongoDB is used as a persistent storage for the back-end to store all the projects
and nodes in the tree structures and also the files uploaded to the system. The Scala logic code
uses a layer called Casbah~\cite{CASBAH} to interface with MongoDB, arrow 4 and 5 in
Figure~\ref{fig:tech_stack}.

\par MongoDB was chosen as the persistent storage because of its quick lookups and because of its
internal storage format called BSON, which is very similar to JSON which the API is using. As the
formats are similar, the process of marshalling and unmarshalling becomes quite easy between the
core code, MongoDB instance and REST interface.  The second reason was that if the system needs to
scale in the future then it is very easy to distribute MongoDB and, if needed, the system can easily
be migrated to Reactive Mongo. Reactive Mongo is an asynchronous and non-blocking driver for MongoDB
and can therefore make the system scale even further~\cite{REACTIVEMONGO}.

\par All files are also stored directly in MongoDB with the help of GridFS. GridFS chunks the files
according to the size limit of MongoDB objects, which is currently 4MB. The advantage of storing
files directly in the database rather than in the filesystem is that backups of can be done through
a database backup which will contain both the meta-data stored in MongoDB and the files which
results in that no separate files needs to be backed up.  Another advantage is that you can retrieve
specific ranges of a file, although that feature is not needed in the Perius implementation. For
scalability this could be used to retrieve different parts of a file from different servers, normal
load balancing would probably work better in a system like Perius where no extremely large files are
expected to be stored.

@@

\end{description}
\subsection{Methods for determining\\implementation details}
This chapter introduces the different methods used to determine how the new system should be
implemented, which persistent storage it should use and how the estimation of long-term scaling 
was done.
\\
\par As this work was not mainly about the system architecture, but rather about evaluating the
effects of using Copy-on-Write, the decisions taken for the different technologies used were simply
made from what the author was comfortable with and expected would have little impact on the
evaluation.  With that said there are no comparisons of for example different databases, protocols
or API solutions in this paper.

@@

\par A low timeout threshold was used to make sure that the system could handle operations faster than they
queued up. This was to be certain that the system could sustain the number of active users during a
longer time frame without having a queue that could grow beyond a manageable size.

@@

\subsubsection{Snapshot Functionality}
The Copy-on-Write concept is not only well-suited for consistency requirements, its basis also gives
the possibility of creating snapshots.

@@

\subsubsection{Examples of Copy-on-Write system implementations}
\subsubsubsection{BTRFS} \label{sec:btrfs}
Btrfs is a B-tree file system for Linux which makes use of Copy-on-Write to make it able to do
efficient writeable snapshots and clones. It also supports cloning of subtrees without having to
actually copy the whole subtree, this is due to the Copy-on-Write effect. As several nodes in the
tree can refer to the same node each node keeps track of how many parents it has by a reference
counter so that the node can be deallocated once the node does not have any parents any more. The
reference counter is not stored in the nodes themselves but rather in a separate data structure so
that a node's counter can be modified without modifying the node itself and therefore avoids the
Copy-on-Write that would have to occur.

@@

\subsubsubsection{Mach kernel}
In the mid 80's when the development of the Mach kernel started, there was problems with that
physically copying memory was too slow. To minimise the copying of memory, Copy-on-Write was
implemented to enable the use of virtual copy operations and to create the possibility of tasks
sharing read-write memory~\cite{MACH}.

\subsubsection{Perius}
As the persistent storage used in this implementation (Section~\ref{sec:tech}) does not implement
transactions or locks, several different problems can occur when multiple clients are working on the
same data set at the same time. Such problems could be race conditions or determining the
happened-before relation~\cite{LAMPORT}. Race conditions are problematic due to that when several
users or nodes modify data they can start in one state and expect the end state of the data to be
the same as if they would have been alone modifying the data, but instead if another user modifies
the data concurrently the changes that the first user made can be overwritten. In this work the
race condition problem is solved by implementing Copy-on-Write on data with high consistency
requirements. The happened-before relation is a problem of ordering events, if it is possible to
know in which order concurrent events happened it is easier to know what the outcome of the events
should be. The happened-before relation is not fully solved in Perius, but suggestions of how to
solve it is given in Section~\ref{sec:conflict_res}.

@@

\subsubsection{Perius}
Perius is the implementation that was done to solve the problem at hand at Uprise. Perius has a
back-end written in Scala and a front-end written in Javascript (ES6), but they are both
independently replaceable. The back-end has a REST API running, which is how the front-end
communicates with the back-end.

@@

The service features a virtual file structure over the assets and metadata that have been stored,
as can be seen in Figure~\ref{fig:frontend}. It also features snapshots, security management of
whole containers as well as individual files, audit and auth logging, multiple separate projects 
and a design which allows for a replaceable persistent storage.

@@

\subsubsection{Entities of the Implementation}
This section explains the different types of entities, and their interaction, that are handled by
Perius. They are explained similarly to the model in Section~\ref{sec:elements}, but with more
implementation-specific details.

@@

In the user interface the stored data is always represented and worked on as a tree, or a forest if
all projects are considered. For a very small project that tree could look like the example in
Figure~\ref{fig:project_tree}.

@@

A very simplified version of how the data is stored can be seen in Figure~\ref{fig:storage}, however
in reality it is not quite so table-like. The implementation is using GUIDs, instead of incremental
numbers as shown in the figure, to work correctly in a distributed environments.

As seen above, the data is not stored as a tree (see Example~\ref{fig:storage}), but rather as
separate documents loosely referring to each other. How the different operations perform their
actions and what implications these actions have on the run-time complexities are not obvious and
that is what is discussed in this section.\\

\begin{figure}[H] 
    \centering
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{|r|c|l|}
        \hline
        \textbf{Variable} & \textbf{Explanation} \\
        \hline
        \textit{n} & All content+containers \\
        \textit{c} & All containers \\
        \textit{m} & All content \\
        \textit{f} & All files \\
        \textit{s} & Size of a file \\
        \textit{k} & Chunk size \\
        \hline
    \end{tabular}
    \renewcommand{\arraystretch}{1.0}
    \caption{Variables used in run-time complexities}
    \label{fig:runtime-variables}
\end{figure}

\par \textbf{GET single entity} \\
When the back-end is handed a GET request it looks at what type of entity that is requested and
returns the document with the correct ID from the corresponding collection. As the collections are
hash indexed on ID this operation can be made in constant time, $O(1)$ and thus $\Theta(1)$. \\

\par \textbf{GET project tree} \\
When requesting a project the full project tree with all entities except files has to be built since
the entities not are stored as a tree in MongoDB. The tree is built layer by layer by fetching
content with the same parent ID, starting from the root of the project. Parent ID's are also indexed
by a hash index which makes it possible to get all nodes with the same parent ID in constant time.
This has to be done for every level with each parent. 
Since all of the containers have to be processed when building the tree, the runtime complexity will
thus be $\Theta(c)$/

\par \textbf{POST/Create entity} \\
Creating a new entity in the system is a constant time operation for containers and content as they
only have to be inserted as new documents in MongoDB, which is a $O(1)$
operation~\cite{MONGOPERFORMANCE}. Even though the insertion itself has constant run-time the index
needs to be updated or sometimes even rebuilt, which can slow down the time before the document is
accessible in constant time. When inserting files they need to be chunked into pieces and inserted
into GridFS, this results in $s/k$ insertions.
The insertion of files can then be regarded as $\Theta(s)$ if the size of the file is regarded, but
if $k$ is larger than the size of the files inserted, the upper bound is constant as the insertion
for files is made in the same way as for other documents.\\

\par \textbf{Snapshot creation} \\
In the creation of a snapshot all metadata in the tree that the snapshot is taken of is duplicated,
meanwhile the files remain the same, this results in $t$ insertions where $t$ is the number of
containers and content entities in the subtree. A snapshots insertion runtime is therefore
$\Theta(n)$. \\

\par \textbf{Delete entity} \\ 
Deleting an entity is a constant time operation when that entity does not have any children. When
the entity has children all those children has to be deleted too which will be done in the same
manner as the project tree is built and therefore results in the same runtime, $\Theta(n)$ for
containers and $\Theta(1)$ for content.

@@

\subsubsection{GET requests} \label{sec:GET_REQUESTS}
To be able to determine what the bottleneck of the back-end is, three different types of \textit{GET
requests} are performed. Most responses are JSON structures (except for the ones in
Listing~\ref{OKREQUEST} and Listing~\ref{STATICREQUEST}) which can be seen in the Response data part
of the listings.\\

@@

\par
The first request (Listing~\ref{OKREQUEST}) tries to maximise the number of requests that spray-can
(HTTP server) can handle with the given specifications and therefore the server just returns 200
(OK). 100 000 requests per second is well over what we need to serve and with a 1.22ms latency we
know that without practically any payload the system is responding fast.

@@

\par
The second request (Listing~\ref{DBREQUEST}) performs the simplest API call which involves getting
all currently stored projects from the MongoDB instance. This request does not yield any extra
computation in the back-end, apart from converting the MongoDB format to the API's JSON result
format. From these requests it can clearly be deduced that the round trip time to fetch a document
in the database is adding some latency, from the average of 1.22ms to 14.52ms with the database
request.

@@

\par
As the second request (Listing~\ref{DBREQUEST}) involves the server answering with some payload that
is generated from the persistent storage, the third request (Listing~\ref{STATICREQUEST}) is simply
the same amount of static payload as the second request but without involving any connections to
MongoDB. This was done to be able to estimate what impact the round trip time to MongoDB and the
conversions to the response format had on the throughput. This test still gave an average latency
that was larger than for the smaller payload in Listing~\ref{OKREQUEST}, but still half of the
latency compared to the request involving a MongoDB call.

@@

The result in Listing~\ref{TREEREQUEST} shows an average of 40 requests/s, which means this is the
real bottleneck in the application. There are two solutions to fix this bottleneck. Since the project
tree will not be modified nearly as often as it is requested, the first solution would be to cache
the results of the tree requests and invalidate those caches when the tree is modified. The second
solution would be to make a more computationally feasible algorithm for building and fetching the whole
tree from the database. This is will not be discussed further, but left as future work. Even
though this massive bottleneck is present, it will not be a problem in a small scale production
environment where the tree is fetched approximately once every 30 seconds by active users, which means
that the application still could support at least 1200 very active users.

@@

\par
With caching turned on, on a single back-end, the application can theoretically support several
million active users that are not posting any content, this is simulated in
Listing~\ref{CACHEREQUEST}.

@@

\par As a project tree is built by querying documents parent ID an optimisation that could be done
is to create an index of the documents parent IDs. This drastically improved the size of the tree
that the server was able to handle within the time-out, going from being able to handle 4000
documents in one tree to at least 50000 documents, which the browser instead will have problems
displaying, represented in the UI, without lag.

\par After adding indices, the building of the project tree was able to finish 50 times faster and
the server was able to serve 2100 requests/s, see Listing~\ref{INDEXREQUEST}. If using the same
approximated average that was used in connection to Listing~\ref{TREEREQUEST} (30 requests/s), the
maximum number of active clients that one node can handle will be >60000 clients at the same time, 
without taking the web server limits or content posting into consideration.
